---
title: "Summer flounder simulation check"
author: "Angelia Miller"
date: "2024-01-04"
output: word_document
---

## Objective 
* simulate 1 year and replicate  



```{r load packages}
# library(stringr)
# library(sf)
# library(patchwork)
library(here)
suppressPackageStartupMessages(library(tidyverse))
library(sdmTMB) 
source(here("R", "StratMeanFXs_v2.R"))

sumflounder.dat <- here("sdmtmb", "sumflounder", "data")
sim.check <- here("sdmtmb", "sumflounder", "data", "simulations", "sim-checking")
```

Model fitting data
```{r}
sf_fall <- readRDS(here(sumflounder.dat, "sumflounder_fall.rds"))
fall_mesh <- readRDS(here(sumflounder.dat, "fall_mesh.rds"))
```

Data wrangle
```{r data}
# read in fall model 
# fall_mod <- readRDS(here(sumflounder.dat, "fall_mod.rds"))
# fall_moddat <- fall_mod$data |> 
#   mutate(EST_YEAR = YEAR, 
#          AREA = as.character(AREA))

# pull polynomial coefficients from fall model fit
fall_x_mat <- poly(sf_fall$AVGDEPTH, 2)
fall_depth_coefs <- attr(fall_x_mat,"coefs")
# saveRDS(fall_x_mat, here(sim.check, "fall_depth_matrix.rds"))
# saveRDS(fall_depth_coefs, here(sim.check, "fall_depth_coefs.rds"))
fall_depth_coefs
```


Re-fit the model 
```{r}
fall_mod <- sdmTMB(EXPCATCHWT ~ poly(AVGDEPTH, 2, coefs = fall_depth_coefs) + 
                    as.factor(EST_YEAR) + 
                    as.factor(AREA) - 1,
                  data = sf_fall,
                  mesh = fall_mesh,
                  family = tweedie(link = "log"), 
                  spatial = "on", 
                  time = "EST_YEAR",
                  spatiotemporal = "IID",
                  control = sdmTMBcontrol(newton_loops = 1)) 

sanity(fall_mod)
#saveRDS(fall_mod, here(sim.check, "fall_mod.rds"))
```


```{r}
coefs <- tidy(fall_mod)
ran_pars <- tidy(fall_mod, "ran_pars")

fall_depth_coefs
coefs
ran_pars
```


Generate predictions
```{r preds}
mod_preds <- predict(fall_mod)
# saveRDS(mod_preds, here(sim.check, "fall_predictions.rds"))

str(mod_preds)
```


Select locations and predictions from one year of past data to simulate values of biomass
```{r resample}
set.seed(1211)
year <- sample(unique(sf_fall$EST_YEAR), size = 1)
# saveRDS(year, here(sim.check, "resample_year.rds"))
resample <- mod_preds |>
  # mutate(EST_YEAR = YEAR) |>
  filter(EST_YEAR == year) |> 
  select(X, Y, STATION, STRATUM, CRUISE6, AVGDEPTH, EST_YEAR, AREA, omega_s) |> 
  mutate(EST_YEAR = 1) |>
  rename(fix_omegas = omega_s)
# saveRDS(resample, here(sim.check, "resampled_data.rds"))
head(resample)

mesh <- make_mesh(resample, xy_cols = c("X", "Y"), cutoff = 10, seed = 42)
# saveRDS(mesh, here(sim.check, "fall_mesh.rds"))
plot(mesh)
```


Simulate without any random effects 
* expected: eta and fix_omegas should match
```{r sim1}
simdat1 <- sdmTMB_simulate(
    formula = ~1 +
      #poly(AVGDEPTH, 2, coefs = fall_depth_coefs) +
      #EST_YEAR +
      #as.factor(AREA) +
      fix_omegas,
    data = resample,
    mesh = mesh,
    family = tweedie(),
    #spatial = "on",
    time = "EST_YEAR",
    #spatiotemporal = "iid",
    B = c(0, 1), 
    range = 0.5, 
    sigma_O = 0,
    sigma_E = 0, 
    phi = 0.1,
    tweedie_p = 1.5,
    seed = 2)
simdat1
# saveRDS(simdat1, here(sim.check, "simdat1.rds"))
```
* result: fix_omegas and eta match. Can add next coefficient. 

Add in a variable and its coeffcient 
```{r sim2}
simdat2 <- sdmTMB_simulate(
    formula = ~1 +
      #poly(AVGDEPTH, 2, coefs = fall_depth_coefs) +
      #EST_YEAR +
      as.factor(AREA) +
      fix_omegas,
    data = resample,
    mesh = mesh,
    family = tweedie(),
    #spatial = "on",
    time = "EST_YEAR",
    #spatiotemporal = "iid",
    B = c(0, 0.7, 1), # added wind coefficient
    range = 0.5, 
    sigma_O = 0,
    sigma_E = 0, 
    phi = 0.1, 
    tweedie_p = 1.5,
    seed = 2)
simdat2

# saveRDS(simdat2, here(sim.check, "simdat2.rds"))
```

Add in 
```{r sim3}
simdat3 <- sdmTMB_simulate(
    formula = ~1 +
      #poly(AVGDEPTH, 2, coefs = fall_depth_coefs) +
      #EST_YEAR +
      as.factor(AREA) +
      fix_omegas,
    data = resample,
    mesh = mesh,
    family = tweedie(),
    #spatial = "on",
    time = "EST_YEAR",
    #spatiotemporal = "iid",
    B = c(0, 0.7, 1), 
    range = ran_pars$estimate[1], # added range parameter
    sigma_O = 0,
    sigma_E = 0, 
    phi = 0.1, 
    tweedie_p = 1.5,
    seed = 2)
simdat3

# saveRDS(simdat3, here(sim.check, "simdat3.rds"))

```


```{r sim4}
simdat4 <- sdmTMB_simulate(
    formula = ~1 +
      #poly(AVGDEPTH, 2, coefs = fall_depth_coefs) +
      #EST_YEAR +
      as.factor(AREA) +
      fix_omegas,
    data = resample,
    mesh = mesh,
    family = tweedie(),
    #spatial = "on",
    time = "EST_YEAR",
    #spatiotemporal = "iid",
    B = c(0, 0.7, 1), 
    range = ran_pars$estimate[1], 
    sigma_O = 0,
    sigma_E = 0, 
    phi = 0.1, 
    tweedie_p = ran_pars$estimate[5], #added power parameter
    seed = 2)
simdat4

# saveRDS(simdat4, here(sim.check, "simdat4.rds"))

```

Add in 
```{r sim5}
simdat5 <- sdmTMB_simulate(
    formula = ~1 +
      #poly(AVGDEPTH, 2, coefs = fall_depth_coefs) +
      #EST_YEAR +
      as.factor(AREA) +
      fix_omegas,
    data = resample,
    mesh = mesh,
    family = tweedie(),
    #spatial = "on",
    time = "EST_YEAR",
    #spatiotemporal = "iid",
    B = c(0, 0.7, 1), 
    range = ran_pars$estimate[1], 
    sigma_O = 0,
    sigma_E = 0, 
    phi = ran_pars$estimate[2], # added phi parameter
    tweedie_p = ran_pars$estimate[5],
    seed = 2)
simdat5

# saveRDS(simdat5, here(sim.check, "simdat5.rds"))
```



Add in 
```{r sim6}
b <- c(0, coefs$estimate[10], 0.7, 1) #save coefficients in object

simdat6 <- sdmTMB_simulate(
    formula = ~1 +
      #poly(AVGDEPTH, 2, coefs = fall_depth_coefs) +
      EST_YEAR + # added year coefficient
      as.factor(AREA) +
      fix_omegas,
    data = resample,
    mesh = mesh,
    family = tweedie(),
    #spatial = "on",
    time = "EST_YEAR",
    spatiotemporal = "IID",
    B = b, # call coefficient here with new value added here
    range = ran_pars$estimate[1], 
    sigma_O = 0,
    sigma_E = 0, 
    phi = ran_pars$estimate[2], 
    tweedie_p = ran_pars$estimate[5],
    seed = 2)
head(simdat6)

# saveRDS(simdat6, here(sim.check, "simdat6.rds"))

simdat6 |>  
  select(!EST_YEAR) |> 
  right_join(mod_preds, by = c("X", "Y")) |> 
  select(X, Y, mu, eta, observed, est, est_non_rf, est_rf, fix_omegas)
```


Here we want `eta` and `est_non_rf` to match in value because we are not simulating random effects 
```{r sim7}
b2 <- c(0, coefs$estimate[1:2], coefs$estimate[10], 0.7, 1) # add poly(depth,2) coeficients

simdat7 <- sdmTMB_simulate(
    formula = ~1 +
      poly(AVGDEPTH, 2, coefs = fall_depth_coefs) + #add depth to predictor
      EST_YEAR +
      as.factor(AREA) +
      fix_omegas,
    data = resample,
    mesh = mesh,
    family = tweedie(),
    #spatial = "on",
    time = "EST_YEAR",
    spatiotemporal = "IID",
    B = b2, 
    range = ran_pars$estimate[1], 
    sigma_O = 0,
    sigma_E = 0, 
    phi = ran_pars$estimate[2], 
    tweedie_p = ran_pars$estimate[5],
    seed = 2)
head(simdat7)

simdat7 |> 
  select(!EST_YEAR) |>
  right_join(mod_preds, by = c("X", "Y")) |> 
  select(X, Y, mu, eta, observed, est, est_non_rf, est_rf, fix_omegas)|> head(n = 10)

# plot(mod_preds$AVGDEPTH, mod_preds$est_non_rf)
# plot(resample$AVGDEPTH, simdat7$`poly(AVGDEPTH, 2, coefs = fall_depth_coefs)1`)
# plot(fall_moddat$AVGDEPTH, fall_x_mat[,1])

```
* at this point is `eta` and `est_non_rf` do not match 

```{r }

```
* adding in the year coefficients and spatiotemporal random effects, drops the `observed` to 0 and `eta` is way lower than the `est` 
